What is an Ancestral Recombination Graph?

### 2022-03-10 paper.tex

Due to the computational expense of sampling ``little ARGs'', state-of-the-art
Monte Carlo methods typically rely on cheaper, approximate models.
The most prominent examples are \texttt{ARGWeaver} \citep{rasmussen2014genome}
and \texttt{Arbores} \citep{heine2018bridging}. The former also utilises a time
discretisation approximation, but scales to dozens of mammal-like genomes.
The \texttt{ClonalOrigin} method \citep{didelot2010inference,
medina2020speeding} uses the \texttt{ClonalFrame}
model and data structure, though it is not as scalable as \texttt{ARGWeaver}.

A very recent MCMC method called \texttt{ARGInfer} has resulted in the first
improvement in the scalability of Monte Carlo methods for the exact ARG model
by making use of the tree sequence representation of marginal trees, suitably enriched
to facilitate the computations needed for MCMC sampling \citep{mahmoudi2021inference}.
While not competitive with the scalability of \texttt{ARGWeaver}, \texttt{ARGInfer} extends
the feasible range of sequence lengths by at least an order of magnitude when
compared with methods which sample realisations of the exact ``little ARG''.

>>>>

Due to the computational expense of sampling ``little ARGs'', state-of-the-art
Monte Carlo methods typically rely on cheaper, approximate models.
The most prominent examples are
\texttt{ARGWeaver}~\citep{rasmussen2014genome,hubisz2020mapping,hubisz2020inference}
and \texttt{Arbores} \citep{heine2018bridging}. The former also utilises a time
discretisation approximation, but scales to dozens of mammal-like genomes.
The \texttt{ClonalOrigin} method \citep{didelot2010inference,
medina2020speeding} uses the \texttt{ClonalFrame}
model and data structure, though it is not as scalable as \texttt{ARGWeaver}.

A very recent MCMC method called \texttt{ARGInfer} has resulted in the first
improvement in the scalability of Monte Carlo methods for the exact ARG model
by making use of the tree sequence representation of marginal trees, suitably enriched
to facilitate the computations needed for MCMC sampling
\citep{mahmoudi2022bayesian}.
While not competitive with the scalability of \texttt{ARGWeaver}, \texttt{ARGInfer} extends
the feasible range of sequence lengths by at least an order of magnitude when
compared with methods which sample realisations of the exact ``little ARG''.
